---
title: "Data Wrangling and Visualization"
author: "Xi Li"
output: pdf_document
---

```{r,include=F}
#knitr::opts_chunk$set(warning = F,error=F,message=F)
require(tidyverse)
#install.packages("dplyr")
require(dplyr)
```




1. Read in and combine all 24 `.csv` into a single data frame in `R` (using a `for()` loop). 
```{r}
#setwd("~/Desktop/AcceleratedStats2/hm1_to_class")
a = list.files("./Data/English-Towns")  # store all file names in this variable
dir = paste("./Data/English-Towns/",a,sep="")
n = length(dir)       
merge.data = read.csv(dir[1])  # read in the first csv
```

```{r}
# read in the rest 23 of csv files
 for (i in 2:n){
   new.data = read.csv(file = dir[i], header=T, sep=",")
   merge.data = rbind(merge.data,new.data)
 }
```

2. Drop the variable `Status`.
3. Change the variable name Ceremonial county to County
4. Order the data by County
5. Export the data as a .csv to the Output/ folder. Title the file “english-town-dictionary.csv”
```{r}
merge.data %>%
  select(-Status) %>%
  # Change the variable name `Ceremonial county` to `County` 
  rename(County=Ceremonial.county) %>%
  # Order the data by `County`
  arrange(County) %>%
  # Save data to the original dataframe
  {. ->> merge.data  } %>%
  # Export the data
  write.csv(.,file = "./Output/english-town-dictionary.csv",row.names=F)
  
```


1. Only countries are in the data; no regions or non-state entities
2. Drop all variables but the `iso3c`,`date`,`value`, `indicatorID`, and `country`.
3. Spread `indicatorID` into two new variables ("SP.POP.TOTL","NY.GDP.MKTP.CD") where the values correspond with the entries in the `value` variable column. 
4. Rename `iso3c` to `ccode`, `date` to `year`, `SP.POP.TOTL` to `pop`, and `NY.GDP.MKTP.CD` to `gdp`.
5. Rearrange the columns in the following order: `country`, `ccode`, `year`, `gdp`, `pop`
```{r}
#install.packages("wbstats")
library(wbstats)
# trying to figure out how it works
#?wbstats
#browseVignettes(package = "wbstats")
```

```{r,cache=TRUE}
# pull data and filter
pop_data <-wb(country="countries_only",indicator=c("SP.POP.TOTL","NY.GDP.MKTP.CD"), startdate = 2000, enddate = 2005)
head(pop_data)
```

```{r}
pop_data %>% 
  # drop useless variables
  select(-indicator,-iso2c) %>% 
  # spread `indicatorID` into two new variables
  spread(indicatorID,value) %>% 
  # rename
  rename(ccode=iso3c,year=date,pop=SP.POP.TOTL,gdp=NY.GDP.MKTP.CD) %>% 
  # rearrange the columns
  select(country,ccode,year,gdp,pop) %>% 
  # save data using popline
  {. ->> pop_data  } 
#pop_data 
```

6. Count to see if you have any missing for `gdp` entries; if there are any missing entries, create a new variable called `missing_gdp` that takes on the value of `1` if missing, `0` otherwise. How many country-year values are missing? Which country has the most missing values?
  - tip: look at the `ifelse()` and `is.na()` functions for ideas regarding how to do this.
  
  3038 country-year values are missing. British Virgin Islands has the most missing values.
```{r}
pop_data$missing_gdp <- ifelse(is.na(pop_data$gdp)==1,1,0)
sum(is.na(pop_data$gdp)) # the number of country-year values that are missing

pop_data %>%
  group_by(country) %>%
  summarise(sum(missing_gdp)) %>%
  arrange(desc(`sum(missing_gdp)`)) # arrange in descending order

```
  
7. Transform `pop` and `gdp` using the natural `log()`. Name the new variables `ln_gdp` and `ln_pop`.
```{r}
pop_data$ln_gdp <- log(pop_data$gdp)
pop_data$ln_pop <- log(pop_data$pop)
#pop_data
```

8. Export the data as a `.dta` and `.csv` into the `Output/` folder. Name the data `wb-pop-gdp-2000-2005`.
```{r}
write.csv(pop_data,file = "./Output/wb-pop-gdp-2000-2005.csv",row.names=F)
library(foreign)
write.dta(pop_data,file = "./Output/wb-pop-gdp-2000-2005.dta")
```


### (1) Use ggplot2 to create a tile plot to demonstrate the portion and location of missing data.

```{r}
# select a part of the data
data <- pop_data[grep("^A",pop_data$country),]
data <- data[as.numeric(data$year)>=2000&as.numeric(data$year)<=2005,]
# convert missing_gpd variable to factor
data$missing_gdp <- as.factor(data$missing_gdp)
```

```{r}
plot1<- ggplot(data=data,aes(x=year,y=country))+
  geom_tile(aes(fill=missing_gdp))+
  theme_minimal()+
  labs(title="Missing GDP Data",
       subtitle = "A visual exploration of data missing GDP in the World Bank sample (2000-2005)",
       caption = "Only country names starting with the letter 'A' included in the figure.",
       x = "Years",
       y = "",
       labels = "Missing")+
  scale_fill_manual(values = c("grey30","Orange"),labels= c("Available","Missing"))+
  theme(legend.position = "top")+
  guides(fill=guide_legend(title="Missing"))

plot1
```

```{r}
# save the plot
ggsave(plot = plot1,filename = "plot1.pdf",
       device = "pdf",width=10,height = 5,path ="./Output")
```

### (2) Generate a scatter plot to present the relationship between GDP and influencing factors.

  
```{r}
data2<- pop_data[as.numeric(pop_data$year)==2000,]
plot2 <- ggplot(data2,aes(x=ln_pop,y=ln_gdp,size = ln_pop))+ # map a continuous variable to size
  geom_point(color = "forestgreen",alpha= 0.5)+
  theme_light()+
  labs(title="Relationship between population and economic development",
       subtitle = "World bank data for the year 2000 reported",
       x = "Log Population",
       y = "Log GDP",
       labels = "Missing")+
  guides(size=FALSE) # drop the legend
  
plot2
```
```{r}
# save the plot
ggsave(plot = plot2,filename = "plot2.png",
       device = "png",width=10,height = 5,dpi = 600,path ="./Output")
```


